---
title: <div align="center">**R PROGRAMMING ASSIGNMENT**</div>
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---
#### _**NAME:** HITESH KISHOR CHAUDHARI_
#### _**PRN:** 210940125013_
<p>&nbsp;</p>
**The dataset has been downloaded from the following link:
<a href = "https://www.kaggle.com/fedesoriano/heart-failure-prediction">_Link_</a>**


```{r echo=F, warning= F, message=F}
library(knitr)
library(kableExtra)
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")
hr <- read.csv("heart.csv",stringsAsFactors = T)

```
<details>
  <summary>Click here to view the Dataset</summary>
    ```{r, echo=FALSE}
    kable(hr,align = "lccrr")%>%kable_styling(font_size = 12,fixed_thead=T)%>%
    scroll_box(height = "500px")
    ```
</details> 
<p>&nbsp;</p>
**1. Considering the categorical variable RestingECG and numerical variable MaxHR, perform one-way ANOVA for means of MaxHR for every category for RestingECG namely, Normal, ST and LVH. Also perform Tukey’s test and write the conclusion of both the hypothesis tests.** 
<p>&nbsp;</p>
```{r echo=FALSE,eval=TRUE}
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")
hdf <- read.csv("heart.csv",stringsAsFactors = T)
anv <- aov(MaxHR~RestingECG,data=hdf)
anova(anv)
TukeyHSD(anv)
```
<details>
  <summary>Click here to view the Conclusion</summary>
<br/>
**Conclusion from ANOVA TEST:-**
As p-value(2.253e-07) < alpha(0.05)
Hence Rejecting H0. We can say means of categories of RestingECG are not same.
At least one of them is different.<br/>

_**Conclusion from TUKEY TEST:-**_
All p-values of Normal,ST,LVH are less than alpha. Hence we can that mean values 
of all categories of RestingECG might be different.
<p>&nbsp;</p>
</details> 


<details>
  <summary>Click here to view the Code</summary>
```{r,echo=T,eval=F}
hdf <- read.csv("heart.csv",stringsAsFactors = T)
anv <- aov(MaxHR~RestingECG,data=hdf)
anova(anv)
TukeyHSD(anv)
```
</details> 

<p>&nbsp;</p>
**2. Calculate mean MaxHR grouped by values in variable RestingECG namely, Normal, ST and LVH. Generate bar graph for the grouped means generated.**
<p>&nbsp;</p>
```{r, echo=FALSE, warning= F, message=F}
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")

library("tidyverse")
hr <- read.csv("heart.csv",stringsAsFactors = T)
hr %>%
  group_by(RestingECG)  %>%
  summarize(MaxHR_mean = mean(MaxHR, na.rm = TRUE))->su
  ggplot(su,aes(RestingECG,MaxHR_mean))+
  geom_bar(aes(fill=MaxHR_mean),stat='identity')+
  labs(title="grouped means plot",x = "RestingECG" ,y = "MaxHR_mean")+
  theme(plot.title = element_text(hjust = 0.5))
```
<details>
  <summary>Click here to view the Code</summary>
```{r,echo=T,eval=F}
library("tidyverse")
hr <- read.csv("heart.csv",stringsAsFactors = T)
hr %>%
  group_by(RestingECG)  %>%
  summarize(MaxHR_mean = mean(MaxHR, na.rm = TRUE))->su
  ggplot(su,aes(RestingECG,MaxHR_mean))+
  geom_bar(aes(fill=MaxHR_mean),stat='identity')+
  labs(title=" Mean MaxHR grouped by RestingECG",x = "RestingECG" ,y = "MaxHR_mean")+
  theme(plot.title = element_text(hjust = 0.5))
```
</details> 
<p>&nbsp;</p>
**3.	Perform two independent samples variance and means tests for value of variable age for two different sex categories namely, Male and Female and write the conclusion of both the hypothesis tests.**
<p>&nbsp;</p>
```{r, echo=FALSE, warning= F, message=F}
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")
df <- read.csv("heart.csv")
male <- df[df$Sex=='M','Age']
female <- df[df$Sex=='F','Age']
var.test(male,female)
t.test(male,female,mu=0,var.equal = T,alternative='l')
```
<details>
  <summary>Click here to view the Conclusion</summary>
<br/>
**Conclusion 1:-**  
As p=0.807 > alpha = .05
There is no difference in variance of age of Male 
and Female, so performing one tail test <br/>

**Conclusion 2:-**
As p=0.9543 > alpha =0.05
accept H0: mu1 >= mu2 is accepted 
Hence we can say age of Male is greater than or equal to Female
 <p>&nbsp;</p>
</details> 

<details>
  <summary>Click here to view the Code</summary>
```{r,echo=T,eval=F}
df <- read.csv("heart.csv")
male <- df[df$Sex=='M','Age']
female <- df[df$Sex=='F','Age']
var.test(male,female)
t.test(male,female,mu=0,var.equal = T,alternative='l')
```
</details> 
<p>&nbsp;</p>
**4.	Calculate mean MaxHR grouped by values in variable HeartDisease namely 1 and 0. Generate bar graph for the grouped means generated.**
<p>&nbsp;</p>
```{r, echo=FALSE, warning= F, message=F}
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")
library("tidyverse")
hdf <- read.csv("heart.csv",stringsAsFactors = T)
hdf %>%
  group_by(HeartDisease)  %>%
  summarize(MaxHR_mean = mean(MaxHR, na.rm = TRUE))->su
#d$mean_size <- factor(d$mean_size)
ggplot(su,aes(HeartDisease,MaxHR_mean))+
  geom_bar(aes(fill=MaxHR_mean),stat='identity')+
  labs(title=" Mean MaxHR grouped by Heart Disease",x = "Heart Disease" ,y = "MaxHR_mean")+
  theme(plot.title = element_text(hjust = 0.5))
```
<details>
  <summary>Click here to view the Code</summary>
```{r,echo=T,eval=F}
library("tidyverse")
hdf <- read.csv("heart.csv",stringsAsFactors = T)
hdf %>%
  group_by(HeartDisease)  %>%
  summarize(MaxHR_mean = mean(MaxHR, na.rm = TRUE))->su)
ggplot(su,aes(HeartDisease,MaxHR_mean))+
  geom_bar(aes(fill=MaxHR_mean),stat='identity')+
  labs(title=" Mean MaxHR grouped by Heart Disease",x = "Heart Disease" ,y = "MaxHR_mean")+
  theme(plot.title = element_text(hjust = 0.5))
```
</details> 
<p>&nbsp;</p>
**5.	Generate scatter plot with X axis as RestingBP and Y axis as MaxHR with the different coloursgiven to the categories of variable HeartDisease. You will have to convert the variable HeartDisease into factor before generating the graph.**
<p>&nbsp;</p>
```{r, echo=FALSE, warning= F, message=F}
setwd("E:/Torrent Downloads/CDAC/R Programming/Datasets")
library("tidyverse")
hdf <- read.csv("heart.csv",stringsAsFactors = T)
hdf$HeartDisease <- factor(hdf$HeartDisease)
ggplot(hdf, aes(RestingBP, MaxHR , color = HeartDisease)) +
  geom_point() +
  labs(title = "RestingBP vs MaxHR") +
  labs(x = "RestingBP", y = "MaxHR ") +
  theme(plot.title = element_text(hjust = 0.5))
```
<details>
  <summary>Click here to view the Code</summary>
```{r,echo=T,eval=F}
library("tidyverse")
hdf <- read.csv("heart.csv",stringsAsFactors = T)
hdf$HeartDisease <- factor(hdf$HeartDisease)
ggplot(hdf, aes(RestingBP, MaxHR , color = HeartDisease)) +
  geom_point() +
  labs(title = "RestingBP vs MaxHR") +
  labs(x = "RestingBP", y = "MaxHR ") +
  theme(plot.title = element_text(hjust = 0.5))
```
</details>
<p>&nbsp;</p>
<p>&nbsp;</p>
<p>&nbsp;</p>





